---
title: "Numbering the strings"
author: "Vivek Trivedi"
date: "26 February 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
```

# Summary

### Not regex

* Use `tidyverse` and `stringr` for messing with strings.
* Massive topic, see the [Resources](http://stat545.com/block028_character-data.html#resources) if a problem occurs. 
* `fruit`, `words`, and `sentences` are prebuilt datasets within `stringr`.
* `str_detect(vector, "str")` : Answers the condition of the object or an element of the object contains the given string, discretely by itself or within a larger element. Also works for regex.
* `append(vector, new_element)` : Adds new elements to a vector.
* `str_subset(vector, "str")` : Similar to `str_detect()`, except this command also produces a new factor with the elements of the given vectors that match the specified condition.
``` {r str-detect}
fruit # chr vector

index <- NULL
item <- NULL

for (n in 1:length(fruit)) {
  if (str_detect(fruit[n], "fruit") == TRUE) {
    index <- append(index, n)
    item <- append(item, fruit[n])
  }
}

(fru <- tibble(index, item))

# OR OR OR...

(fru2 <- str_subset(fruit, "fruit")) # It would be nice to have indices though.
```

* `str_split(vector, "delimiter")` : Outputs a list of elements of a vector split by the given delimiter. Often the delimiter is space, tab, comma, etc. 
* `str_split_fixed(vector, "delimiter" n = num)` : Outputs a length(vector) x num matrix of elements split by the given delimiter. 
* `separate(tbl_df, col, into = c("col1, col2"), sep = "bla")` : Same as `str_split()`, but works on data frames instead of vectors. Produces a tidy data frame with the given column names. **Note:** `into` argument needs to be supplied. It has no default.
``` {r splitting}
str_split(fru2, " ") # class list
str_split_fixed(fru2, " ", n = 3) # class matrix
separate(fru, item, c("x" , "fruit"), " ") # class tbl_df
```

* `str_length(vector)` : Gives the length of strings within a vector.
* `str_sub(vector, start_pos, end_pos)` : Output substrings including and in between the given character positions of the vector elements. Sliding windows and assignments are also possible (with caveats). Also see `str_replace()`.
``` {r length&sub}
str_length(fru2) # vector
str_sub(fru2, 4, ) # vector
str_sub(fru2, 1:6, 3:8) # vector
fru3 <- fru2
mehfruit <- str_sub(fru3, 1, 4) <- "meh" 
mehfruit # Not how it works...
# Only works this way...
str_sub(fru3, 1, 4) <- "meh"
fru3 # vector
```

* `str_c(vector_elements1, vector_elements2, sep = "bli", collapse = "bla")` : Useful little thing that can concantenate strings in a vector into larger strings. Opposite of `str_split()` and `separate()`.
* `unite(tbl_df, "new_col", col1, col2..., sep = " or ")`: Combines two character-type tibble columns into a single column with a combiner string.  
``` {r concantenate-collapse}
str_c(fru2[1:4], fru2[5:8], sep = " or may be ")
str_c(fru2, collapse = ", ")
str_c(fru2[1:4], fru2[5:8], sep = " or may be ", collapse = ", ")
fruco <- tibble(fruit1 = fru$item[1:4], fruit2 = fru$item[5:8]) # unite() doesn't like col arguments with indices, therefore, simplifying column names is necessary.
fruco %>%
   unite("fruits", fruit1, fruit2, sep = " & then ")
```

* `str_replace(vector, "replacee", "replacer")` : Replaces a substring with another string. `str_replace_na()` variant exists.
* `replace_na(tbl_df, replace = list(col = "replacer"))` : Replaces `NA` values in a tibble with the replacer string.
``` {r replace} 
fru21 <- fru
str_replace(fru21$item, "uit", "actal")
fru21$item[3] <- NA # NULL doesn't work for some reason...
fru21$item
str_replace_na(fru21$item, "It works") # returns vector.
replace_na(fru21, replace = list(item = "NADA")) # returns data frame.
```

---

### Regex

* It's like learning a new language no one even likes to speak or read...
* When strings cannot be called upon using another fixed string (e.g.: `str_replace(fru, "fruit", "vege")`), regular expressions become useful, as they deal with character patterns within strings.
* Regex go between quoation marks.
* When using `\`, make sure to put an extra `\` to escape the stringification of stuff between quotation marks. Double square brackets around `:bla:` for the same reason. 

Regex|  Meaning
-----|----------
`.`  | a single character
`\n` | new line
`^`  | beginning of string
`$`  | end of string
`\b` | word boundary
`\B` | not word boundary
`\d` | single digit
`/s` | space aka `[:space:]`
`[bla]` | character class
`[^bla]` | not bla chr class
`[:punct:]` | Punctuation
`*` | 0 or more
`+` | 1 or more
`?` | 0 or 1
`{n}` | Exactly n
`{n,}` | at least n
`{,m}` | Up to m
`{n,m}` | Between and including n and m



``` {r regex}

```

