---
title: "Tidyin' the Mess"
author: "Vivek Trivedi"
date: "6 February 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
# Summary
* R is a data-aware language. It instrinsically knows the structure of data.
* Tidy data is (essentially) a single data frame with columns representing variables and rows representing observations. It looks not very pretty to us, but R loves tidy data data aggregation for graphics.
* Remember, the first argument is most of these functions is the initial object to be manipulated. Therefore, use the pipe operator if needed.

---

### Operations for similar data frames
* `read_csv(file.path(path/to/file))`: Imports csv files into the workspace. Other `read` variants [also exist](https://github.com/tidyverse/readr). `read_csv(chrDataFrame, trim_ws = TRUE, skip = 1)` can extract comma separated data from a character type object into a tibble/data frame.
``` {r import, message = FALSE}
# library(tidyverse)
males <- read_csv(file.path("males.csv"))
females <- read_csv(file.path("females.csv"))
```

* Use `message = FALSE` in RMarkdown code chunk headers to hide code execution messages from being displayed.
* `bind_rows(df1, df2...)`: Combines multiple dataframes into a single one by putting the names of the dataframes in a separate column/variable.
``` {r bind}
data_untidy <- bind_rows(males, females)
data_untidy
```

* `gather(untidy_df, key = 'new_col/var', value = 'numbers', old_col1, old_col2...)`: Turns the title of column/s in a dataframe into an observation in each row/sample. 
``` {r gather}
data_tidy <- gather(data_untidy, key = 'Species', value = 'Words', Elf, Hobbit, Man)
data_tidy
```

* `write_csv(tidy_df, file.path("desired/path/to/csv"))`: Write dataframes to given location as a .csv file.
``` {r write}
write_csv(data_tidy, file.path("lotr_tidy.csv"))
```

* `unite(tib_df, combined_var, var1, var2)`: Combines the data and names of two columns into a single column.
``` {r unite}
data_mashed <- data_tidy %>%
  unite(Race_Gender, Species, Gender)

data_mashed
```

* `spread(tidy_df, key = new_col/vars, value = old_col/vars)` : Opposite of `gather()`. Turns the observations of the `key` variable into column/variable. The result is much nicer to human eyeballs.
``` {r spread}
data_mashed %>%
  spread(key = Race_Gender, value = Words) %>%
  print(width = Inf)
```

---

### Operations for (kind of) disparate data frames
* `read_csv(chrDataFrame, trim_ws = TRUE, skip = 1)`: Extracts comma separated data from a character type object into a tibble/data frame. The `trim_ws` argument is used to exclude the preceding and succeeding whitespace around the data. `skip` is used to point out how many lines from the top are not part of the data.
``` {r superhero-entry}
superheroes <- "
name,   alignment,  gender, publisher   
Magneto,   bad,    male,   Marvel
Storm,  good,   female, Marvel
Mystique,   bad,    female,    Marvel
Batman,   good,   male,   DC
Joker,  bad,    male,   DC
Catwoman,   bad,    female, DC
Hellboy,    good,   male,   Dark Horse Comics
" # Data separated by one tab. The next action works with other types of whitespaces too (such as space).

superheroes <- read_csv(superheroes, trim_ws = TRUE, skip = 1)

publishers <- "
  publisher, yr_founded
         DC,       1934
     Marvel,       1939
      Image,       1992
"
publishers <- read_csv(publishers, trim_ws = TRUE, skip = 1)

superheroes
publishers
```

* `inner_join(df1, df2)`: Kind of like df1 intersect df2. Returns all rows from x where there are matching values in y, and all columns from x and y.
``` {r inner-join}
inner_join(superheroes, publishers)
```
Notice how Hellboy is not present in the dataframe, as the "publisher" value of that row is not present in the `publishers` data frame. Also, the publisher "Image" is not present in the result due to the same reason.

* `semi_join(df1, df2`: Return all rows from x where there are matching values in y, keeping just columns from x. A semi join differs from an inner join because an inner join will return one row of x for each matching row of y, where a semi join will never duplicate rows of x. This is a filtering join.
 