---
title: "Tidyin' the Mess"
author: "Vivek Trivedi"
date: "6 February 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gapminder)
g <- gapminder
```
# Summary
* R is a data-aware language. It instrinsically knows the structure of data.
* Tidy data is (essentially) a single data frame with columns representing variables and rows representing observations. It looks not very pretty to us, but R loves tidy data data aggregation for graphics.
* Remember, the first argument is most of these functions is the initial object to be manipulated. Therefore, use the pipe operator if needed.
* Be vary of commands that change the arrangement/order of rows.

---

### Operations for similar data frames
* `read_csv(file.path(path/to/file))`: Imports csv files into the workspace. Other `read` variants [also exist](https://github.com/tidyverse/readr). `read_csv(chrDataFrame, trim_ws = TRUE, skip = 1)` can extract comma separated data from a character type object into a tibble/data frame.
``` {r import, message = FALSE}
# library(tidyverse)
males <- read_csv(file.path("males.csv"))
females <- read_csv(file.path("females.csv"))
```

* Use `message = FALSE` in RMarkdown code chunk headers to hide code execution messages from being displayed.
* `bind_rows(df1, df2...)`: Combines multiple dataframes into a single one by putting the names of the dataframes in a separate column/variable.
``` {r bind}
data_untidy <- bind_rows(males, females)
data_untidy
```

* `gather(untidy_df, key = 'new_col/var', value = 'numbers', old_col1, old_col2...)`: Turns the title of column/s in a dataframe into an observation in each row/sample. 
``` {r gather}
data_tidy <- gather(data_untidy, key = 'Species', value = 'Words', Elf, Hobbit, Man)
data_tidy
```

* `write_csv(tidy_df, file.path("desired/path/to/csv"))`: Write dataframes to given location as a .csv file.
``` {r write}
write_csv(data_tidy, file.path("lotr_tidy.csv"))
```

* `unite(tib_df, combined_var, var1, var2)`: Combines the data and names of two columns into a single column.
``` {r unite}
data_mashed <- data_tidy %>%
  unite(Race_Gender, Species, Gender)

data_mashed
```

* `spread(tidy_df, key = new_col/vars, value = old_col/vars)` : Opposite of `gather()`. Turns the observations of the `key` variable into column/variable. The result is much nicer to human eyeballs.
``` {r spread}
data_mashed %>%
  spread(key = Race_Gender, value = Words) %>%
  print(width = Inf)
```

---

### Operations for (kind of) disparate data frames
* **Caution:** Join functions love changing the order of the original data. Arrange, filter and ascertain data order appropriately before performing analysis.
* **Also, caution:** Order of data frames in arguments matters. A lot.
* `read_csv(chrDataFrame, trim_ws = TRUE, skip = 1)`: Extracts comma separated data from a character type object into a tibble/data frame. The `trim_ws` argument is used to exclude the preceding and succeeding whitespace around the data. `skip` is used to point out how many lines from the top are not part of the data.
``` {r superhero-entry}
superheroes <- "
name,   alignment,  gender, publisher   
Magneto,   bad,    male,   Marvel
Storm,  good,   female, Marvel
Mystique,   bad,    female,    Marvel
Batman,   good,   male,   DC
Joker,  bad,    male,   DC
Catwoman,   bad,    female, DC
Hellboy,    good,   male,   Dark Horse Comics
" # Data separated by one tab. The next action works with other types of whitespaces too (such as space).

superheroes <- read_csv(superheroes, trim_ws = TRUE, skip = 1)

publishers <- "
  publisher, yr_founded
         DC,       1934
     Marvel,       1939
      Image,       1992
"
publishers <- read_csv(publishers, trim_ws = TRUE, skip = 1)

superheroes
publishers
```

**Going from most mutating to most filtering...**

* `full_join(df1, df2)` : df1 union df2. Returns all rows and columns for df1 and df2, but with `NA` for missing values.
``` {r full-join}
full_join(publishers, superheroes)
```

* `left_join(df1, df2)`: Basically, df1 union df2, except only returns the rows from df1. Multiple matches produce multiple rows. Puts down `NA` values for non-matching values.
``` {r l-join}
left_join(publishers, superheroes)
```
Notice that the Hellboy row from `superheroes` does not appear in the result, as it has not matching value in df1, `publishers`.

* `inner_join(df1, df2)`: Kind of like df1 intersect df2. Returns all rows from df1 where there are matching values in df2, and all columns from df1 and df2.
``` {r inner-join}
inner_join(superheroes, publishers)
```
Notice how Hellboy is not present in the dataframe, as the "publisher" value of that row is not present in the `publishers` data frame. Also, the publisher "Image" is not present in the result due to the same reason.

* `semi_join(df1, df2)`: Sort of `inner_join()`, except only returns rows from df1. Filters and selects df1 rows with matching values in df2.
``` {r semi-join}
semi_join(publishers, superheroes)
```

* `anti_join(df1, df2)` : df1 - (df1 intersect df2), but only with the columns of df1. If a row in df1 has a matching value in df2, it will be filtered out of the results. Opposite of `semi_join()`.
``` {r anti-join}
anti_join(superheroes, publishers)
```

* `tibble(col1 = {bla}, col2 = {blabla}...)` : creates a tibble data frame with each argument and its equivalent defining the columns. Use `tribble()` to enter data by rows.
* `droplevels()` : Drops all unused levels from a tbl_df. No idea why it's really necessary...
``` {r tribble}
lookupDf <- tibble(country = c("Belgium", "India", "United States", "Canada"), food = c("waffle", "Tikka", "Twinkie", "poutine"))

# OR OR OR...

lookupDf <- tribble(
  ~ country, ~food,
  "Belgium",  "waffle",
  "India", "Tikka",
  "United States", "Twinkie",
  "Canada", "poutine"
)

# Just for fun...

lookupGm <- g %>%
  filter(country %in% c("Belgium", "India", "United States", "Canada", "Malawi"), year > 2003) %>%
  select(-pop, -gdpPercap) %>%
  droplevels()

str(lookupGm)
```