---
title: 'Demo: Messin'' with dataframes'
author: "Vivek Trivedi"
date: "5 February 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gapminder)
library(tidyverse)
```

# Investigative question...
**Was there ever a country that did not have a correlational relationship between GDP per capita and life expectancy in a five year period between 1955 and 2007? If so, why did that occur? If not, why did this not occur in recent memory?**

# Process
Set up environment `-->` Data `-->` ascertain its order and arrangement `-->` group by continent, country `-->` add columns for change in GDP Per Capita and life expectency from previous five years `-->` verify `-->` make a new bare-bones copy of data `-->` filter for lack of coorelational relationship between the two `-->` ??? `-->` Profit.

#The Magic
``` {r prelim}
rm(list=ls())
graphics.off()

library(gapminder)
library(tidyverse)

g <- gapminder
```

``` {r part-0}
analysis_1 <- g %>%
  arrange(country, year) %>%
  group_by(continent, country) %>%
  mutate(changeGdpPC = ((gdpPercap - lag(gdpPercap))*100/lag(gdpPercap)),
         changeLEPC = ((lifeExp - lag(lifeExp))*100/lag(lifeExp))) # Brackets galore

options(tibble.width = Inf) # Shows all columns of a tibble... a bit ugly though.
analysis_1
```
**Looking alright so far...**


# Countries with a five yearly increase in GDP Per Capita and a decrease in life expectancy...
``` {r gdp-up-exp-down}
finale_data <- analysis_1 %>%
  select(country, continent, year, changeGdpPC, changeLEPC)

prod_1 <- finale_data %>%
  filter(changeGdpPC > 0) %>%
  filter(changeLEPC < -1) %>% # Arbitary threshold to reduce noise
  arrange(changeLEPC)

print(prod_1, n = 50)
```

# Countries with a five yearly increase in Life Expectancy and an increase in GDP Per Capita...
``` {r gdp-down-lifeExp-up}
finale_data <- analysis_1 %>%
  select(country, continent, year, changeGdpPC, changeLEPC)

prod_2 <- finale_data %>%
  filter(changeGdpPC < -5) %>% # Arbitary thresholds to reduce noise
  filter(changeLEPC > 1) %>%
  arrange(changeGdpPC)

print(prod_2, n = 50)
```


# Fishing for interesting crumbs...
```{r plot-1}
plot(prod_1$changeGdpPC~prod_1$changeLEPC)
hist(prod_1$year)
plot(prod_1$continent) 
# hist() did not work for some reason, eventhough the continents
# are levels within the factor "continent".
```

```{r plot-2}
plot(prod_2$changeLEPC~prod_2$changeGdpPC)
hist(prod_2$year)
plot(prod_2$continent)
```

